---
title: "Discrete Random Variables"
output:
  html_notebook: default
  html_document: default
---

# Random Variable
A random variable is a function which assign a value from the sample space to real numbers. It relates the value of the probability associated with an event in the sample space.


## Bernoulli Random Variable

Consider the tos of a coin, which comes up a head with probability *p*, and a tail with probability *1-p*. The *Bernoulli* random variable takes the two values 1 and 0, depending on whether the outcome is a head or a tail:


$$ X = {1, head \brace 0, tail} $$ 
Its PMF is,

$$ p_{X}{k} = {p, k=1 \brace 1-p, k=0}  $$


## Binomial Random Variable

A coin is tossed n times. At each toss, the coin comes up a head with probability *p*, and a tail with probability *1-p*, independent of prior tosses. Let *X* the number of heads in the n-toss sequence. We refer to *X* as a *binomial* random variable with parameters *n* and *p*. The *PMF* of *X* consist of the binomial probabilities that were:


$$p_{X}(k) = P(X = k)= \binom{n}{k}p^{k}(1-p)^{n-k}$$
Donde $$ k = 0,1,...,n $$

### Problem 1.1

In 10 trials with success probability 0.5, what is the probability that no more than 4 success wil be observed?

### Solution

Here, n=10, p = 0.5, and we want to calculate

$$ P(Y<=4) = F(4)  $$

We do so in R as follows:

```{r,eval=TRUE}
res <- pbinom(q = 4, size= 10, prob = 0.5)
res
```


### Example 1.2


In 12 trials with success probability 0.3, what is the probability that more than 6 successes will be observed?

### Solution

Here, n = 12, prob=0.3 and P(6<Y).

Since $P(6<Y)$, this can be traslated  $P(6<Y) = 1 - P(Y<=6) = 1 - F(6).$

We do this so in R as follows:

```{r,eval=TRUE}
res <- 1-pbinom(q = 6 , size= 12, prob = 0.3)
res
```

### Problema 1.3

In 15 trials with success probability 0.6, what is the probability that at least 5 but no more than 10 successes will be observed?

### Solution

Here n = 15, prob=0.6, and we want to calculate:

$P(5=<Y<=10)$, this can be traslated to $P(5<=Y<=10) = P(X<=10) - P(x>4)= F(10)-F(4)$

We do so in R as follows:

```{r,eval=TRUE}
res <- pbinom(10, size = 15, prob = 0.6) - pbinom(4, size = 15, prob = 0.6) 
res
```


### Problem 1.4

In 20 trials with success probability 0.9, what is the probability that exactly 16 successes will be observed?

### Solution

$P(X=16)=P(Y<=16)-P(Y<=15)=F(16)-F(15)$


We do so in R as follows:

```{r, eval=TRUE}
res <- pbinom(16, size = 20, prob = 0.9) - pbinom(15, size = 20, prob=0.9)
res
```








### Problem 1.5

Suppose there are twelve multiple choice questions in an English class quiz. Each question has five possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.

### Solution

El espacio muestral es *n = 12* (twelve questions). Dado que son 5 alternativas para cada pregunta, la posibilidad de responder de forma correcta (y aleatoria) es *1/5=0.2*.

Queremos encontrar 4 o menos (P(X<=4)) respuestas correctas:
P(X=4) + P(X=3) + P(X=2) + P(X=1) + P(X=0). Es decir, tener 4 respuesta correctas, 3 respuestas correctas, 2 respuestas correctas,..., 0 o ninguna respuesta correcta. 

Calculemos primero para 4 respuestas correctas:
dbinom(x = 4, p = 0.2, n = 12)

```{r, eval=TRUE}
x <- dbinom(x = 4, p = 0.2, size = 12)
x
```

Calculemos el resto de probabilidades:

```{r, eval=TRUE}
cum_x <- dbinom(0, size=12, prob=0.2) + dbinom(1, size=12, prob=0.2) + dbinom(2, size=12, prob=0.2) + dbinom(3, size=12, prob=0.2) + dbinom(4, size=12, prob=0.2) 
cum_x
```

Alternativamente 

```{r,eval=TRUE}
cum_x <- pbinom(q = 4, size= 12, prob = 0.2)
cum_x
```

Por lo tanto la posibilidad de responder 4 o menos preguntas correctamente de forma aleatoria un cuesteonario de 12 preguntas es *0.9274445*

El grafico de la distribucion binomial

```{r,eval=TRUE}
case = c(0,1,2,3,4)
point = c(0.06871948, 0.2061584, 0.2834678, 0.2362232, 0.1328756)
dat <- data.frame(case, point, stringsAsFactors = FALSE)
plot(dat$case, dat$point, type = "h")
```















