---
title: "Discrete Random Variables"
output:
  html_notebook: default
  html_document: default
---

# Random Variable
A random variable is a function which assign a value from the sample space to real numbers. It relates the value of the probability associated with an event in the sample space.

## Probability Mass Function (PMF)

In simple terms, the *Probability Mass Function (PMF)* is the probability of a random variable takes a value *x*. 

Mathematically, the *Probability Mass Function (PMF)* is a function that assigns the probability $p_{X}$ of the event *{X=x}* consisting of all *outcomes* (the outcomes are denoted by $\omega$ , $\omega$  belongs to the sample space  $\Omega$), such that $X(w) = x$) that give rise to value of *X* equal to *x*: 

The *PMF* of *x* denoted by $p_{X}(x)= P( \lbrace{X=x}\rbrace )=P(X=x)$



## Bernoulli Random Variable

Consider the tos of a coin, which comes up a head with probability *p*, and a tail with probability *1-p*. The *Bernoulli* random variable takes the two values 1 and 0, depending on whether the outcome is a head or a tail:


$$ X = {1, head \brace 0, tail} $$ 
Its PMF is,

$$ p_{X}{k} = {p, k=1 \brace 1-p, k=0}  $$


## Binomial Random Variable

A coin is tossed n times. At each toss, the coin comes up a head with probability *p*, and a tail with probability *1-p*, independent of prior tosses. Let *X* the number of heads in the n-toss sequence. We refer to *X* as a *binomial* random variable with parameters *n* and *p*. The *PMF* of *X* consist of the binomial probabilities that were:


$$p_{X}(k) = P(X = k)= \binom{n}{k}p^{k}(1-p)^{n-k}$$
Donde $$ k = 0,1,...,n $$

### Problem 1.1

In 10 trials with success probability 0.5, what is the probability that no more than 4 success wil be observed?

### Solution

Here, n=10, p = 0.5, and we want to calculate

$$ P(Y<=4) = F(4)  $$

We do so in R as follows:

```{r,eval=TRUE}
res <- pbinom(q = 4, size= 10, prob = 0.5)
res
```


### Example 1.2


In 12 trials with success probability 0.3, what is the probability that more than 6 successes will be observed?

### Solution

Here, n = 12, prob=0.3 and P(6<Y).

Since $P(6<Y)$, this can be traslated  $P(6<Y) = 1 - P(Y<=6) = 1 - F(6).$

We do this so in R as follows:

```{r,eval=TRUE}
res <- 1-pbinom(q = 6 , size= 12, prob = 0.3)
res
```

### Problema 1.3

In 15 trials with success probability 0.6, what is the probability that at least 5 but no more than 10 successes will be observed?

### Solution

Here n = 15, prob=0.6, and we want to calculate:

$P(5=<Y<=10)$, this can be traslated to $P(5<=Y<=10) = P(X<=10) - P(x>4)= F(10)-F(4)$

We do so in R as follows:

```{r,eval=TRUE}
res <- pbinom(10, size = 15, prob = 0.6) - pbinom(4, size = 15, prob = 0.6) 
res
```


### Problem 1.4

In 20 trials with success probability 0.9, what is the probability that exactly 16 successes will be observed?

### Solution

$P(X=16)=P(Y<=16)-P(Y<=15)=F(16)-F(15)$


We do so in R as follows:

```{r, eval=TRUE}
res <- pbinom(16, size = 20, prob = 0.9) - pbinom(15, size = 20, prob=0.9)
res
```



### Problem 1.5

Suppose there are twelve multiple choice questions in an English class quiz. Each question has five possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.

### Solution

El espacio muestral es *n = 12* (twelve questions). Dado que son 5 alternativas para cada pregunta, la posibilidad de responder de forma correcta (y aleatoria) es *1/5=0.2*.

Queremos encontrar 4 o menos (P(X<=4)) respuestas correctas:
P(X=4) + P(X=3) + P(X=2) + P(X=1) + P(X=0). Es decir, tener 4 respuesta correctas, 3 respuestas correctas, 2 respuestas correctas,..., 0 o ninguna respuesta correcta. 

Calculemos primero para 4 respuestas correctas:
dbinom(x = 4, p = 0.2, n = 12)

```{r, eval=TRUE}
x <- dbinom(x = 4, p = 0.2, size = 12)
x
```

Calculemos el resto de probabilidades:

```{r, eval=TRUE}
cum_x <- dbinom(0, size=12, prob=0.2) + dbinom(1, size=12, prob=0.2) + dbinom(2, size=12, prob=0.2) + dbinom(3, size=12, prob=0.2) + dbinom(4, size=12, prob=0.2) 
cum_x
```

Alternativamente 

```{r,eval=TRUE}
cum_x <- pbinom(q = 4, size= 12, prob = 0.2)
cum_x
```

Por lo tanto la posibilidad de responder 4 o menos preguntas correctamente de forma aleatoria un cuesteonario de 12 preguntas es *0.9274445*. El grafico de la distribucion binomial:

```{r , fig.align='center'}
case = c(0,1,2,3,4)
point = c(0.06871948, 0.2061584, 0.2834678, 0.2362232, 0.1328756)
dat <- data.frame(case, point, stringsAsFactors = FALSE)
plot(dat$case, dat$point, type = "h")
```


## The Geometric Random Variable

We can interpret the geometric random variable in terms of repeated independent trials until the first *success*. Each trial has probabilit of success *p* and the number of trials until (and including) the first success is modeled by the geometric random variable. The meaning of *success* is context-dependent. For example:

- Finding a missing item in a given search.
- Passing a test in a given try.
- Finding a adn sequence given an experiment.


Suppose that we repeatedly and independently toss a coin with probability of a head equal to p, where *0<p<1*. The *geometric random variable* is the number of $X$ of tosses for a head to come up for the first time. Its PMF is given by:

$p_{X}(k) = (1-p)^{k-1}p$, where $k=1,2,3,4.....$



## Poisson Distribution

From prior data we can estimate the average number of events per unit of time or space, but we might also want to know how different this might be from one unit of time/space to another. The Poisson distribution tells us the distribution of events per unit of time or space, when we sample many such units. It is useful when addressing queuing questions like **how much capacity do we need to be 95% sure of fully processing the internet traffic that arrives on a server in any 5 second period.** [2].


Mathematically, a random variable $X$, taking on one of the values $0, 1, 2, ... ,$ is said to be a $Poisson Random Variable$ with parameter $\lambda$, if for some $\lambda >0$, the P.M.F is:


$P_{X}(k)= P{(X=k)} = e^{-\lambda}\frac{\lambda^{k}}{k!}$


The key parameter in a Poisson distribution is $\lambda$, or lambda. This is the expected value ($E(X)$) or mean number of events that occurs in a specified interval of time or space. The variance ($Var(X)$) for a Poisson distribution is also $\lambda$.


## R implementation of Poisson Random Variable

A common technique is to generate random numbers from a Poisson distribution as part of a queuing simulation. The `rpois` function in R does this, taking only two arguments - the quantity of random numbers sought, and lambda:

```{r}
#Poisson random numbers
prn <- rpois(100, lambda = 2)
prn
```
```{r}
hist(prn)
```

This code will generate $100$ random numbers from a Poisson distribution with $λ = 2$. For example, if incoming customer service calls average two per minute, this code will simulate $100$ minutes, returning the number of calls in each of those $100$ minutes.

### Problem 1.6

The average number of beetle bites by seed is 0.2 (i.e, for instance, in average each 100 seeds you count 20 bites). Questions: 
- How many seeds would not have bites if we evaluate 100 seeds?  
- How many with 1 bites? and How many 2 or more?

### Solution

To answer this questions, we will use the Poisson Distribution as follow:

-If there is no bites ($X=0$) and the average ($\lambda=0.2$), the P.M.F is:

$P{(X=k)} = e^{-\lambda}\frac{\lambda^{k}}{k!}$
$P{(X=0)} = e^{-0.2}\frac{0.2^{0}}{0!}$
$P{(X=0)} = 0.819$ 

In *R*:

```{r}
dpois(0, 0.2, log = FALSE)
```

-If there is one bite per seed ($X=1$) and the average ($\lambda=0.2$), the P.M.F is:

$P{(X=k)} = e^{-\lambda}\frac{\lambda^{k}}{k!}$
$P{(X=1)} = e^{-0.2}\frac{0.2^{1}}{1!}$
$P{(X=1)} = 0.164$

In *R*:

```{r}
dpois(1, 0.2, log = FALSE)
```

- If exist 2 or more bites per seed,

$P{(X>k)} = P(X=k-1)+ P(X=k-2)$
$P{X>1} = P(X=1)+ P(X=0)$
$P{X=1} = 0.164 + 0.819 = 0.018$

- How many seeds would not have bites if we evaluate 100 seeds?

We obtained that for $X=0$ (no bites) we have a probabilitt of $p=0.819$ . So for 100 seed we aproximately have *(100)x(0.819)= 82 no bites*

- In 100 seed, find 1 bite is aproximately:  $(100x0.164)= 16$

- In 100 seed, find 2 bite or more is aproximately:  $(100x0.018)= 2$


# Bibliography

[1] Introduction to Probability-Dimitri P. Bertsekas, John N. Tsitsiklis  2nd Edition

[2] Practical Statistics for Data Scientists 50 Essential Concepts Peter Bruce, Andrew Bruce 

[3] Basic Probability Distributions in R https://rstudio-pubs-static.s3.amazonaws.com/100906_8e3a32dd11c14b839468db756cee7400.html
